Refinement of Thermostated Molecular Dynamics Using Backward Error Analysis

### 2019-02-26 response_to_reviewers.tex

{\color{blue}{
- p.8, middle of left column: ``which differs from the others for involving a global stochastic thermostat", should be ``by involving". 
}}

\textbf{Answer}:
Done.

{\color{blue}{
- p.8, right column, paragraph starting with ``Fig. 5(d) contains...". The conclusion in this paragraph seems dubious. The authors appear to confuse numerical artifacts, due to finite time step, with the physical phenomenon of heat capacity dependence on temperature. A much more plausible explanation is that statistical error in the measurement of $c_v$ (which is related to the variance of the total energy) is much larger (for the same simulation sample) than that in other quantities (like kinetic or potential energy). So, the statistical error masks the systematic error due to finite time step. In other words, the dependence of measured $c_v$ on time step size is present, but would require more simulation time in order to measure with sufficient precision. 
}}

\textbf{Answer}:
Our (unfulfilled) intention here was to suggest that the effect caused by numerical artifact is similar to that associated with small temperature variations in experiments, where the average energy changes while $c_v$ remains practically unaltered.
As the reviewer pointed out, $c_v$ is related to the variance of the total energy distribution.
However, the results show that such variance is not being affected by discretization errors in the same extent as the average is.
To emphasize this point, we have rewritten the paragraph while avoiding the analogy with the physical phenomenon.
We have also employed the expression ``are not noticeably influenced'' in place of ``does not get affected'', in order to acknowledge the statistical nature of these measurements.

@@

{\color{blue}{
- Fig. 5 end of caption: replace ``consists in'' with ``is", as in ``Finally, the KLN scheme (triangle) is the double-thermostat method...". 
}}

\textbf{Answer}:
Done.

{\color{blue}{
- p.10, bottom of left column: ``Having overcome such barrier ...'' I'm not sure what ``barrier'' the authors have in mind. What they propose allows to increase the accuracy of the results (maybe double it, if we are talking about increasing the time step from 1fs to 4fs), or get the same accuracy with 25\% of the computational resource (again from 1fs to 4fs), but this is not overcoming a ``barrier". So, I suggest either deleting this sentence as unnecessary, or modifying it to reflect more accurately the contribution of this work. 
}}

\textbf{Answer}:
%We did not have anything unusual in mind.
This was simply an unfortunate choice of words.
We deleted the sentence ``Having overcome such barrier'' and modified the paragraph so as to describe our contribution in a simpler way.

@@

\section{\textbf{Reviewer \#2}}

{\color{blue}{
The manuscript presents results from a computational study examining the influence of integration errors on kinetic energy equipartitioning in canonical ensemble MD simulations. The authors argue that an auxiliary shadow kinetic energy should be used instead of the conventional definition when formulating thermostats and analyzing equipartitioning behavior. They show that use of this shadow kinetic energy leads to improved numerical stability and subsequently demonstrate that accurate estimates of properties in the target ensemble can be obtained by reweighting. This reformulation approach is shown to work successfully for several widely-used thermostats.

This is a nicely presented article that is likely to be of interest the readership of JCP. Overall, it is technically sound, but I have a couple of comments that the authors should consider. 
}}

\textbf{Answer}:
 We sincerely thank the reviewer for the kind comments and helpful suggestions. They have been considered in full.

### 2019-02-26 rigid_bodies.tex
Small changes

Fig.~\ref{fig:properties}(d) contains results for the constant-volume specific heat capacity, which
are not noticeably influenced by the discretization errors when $\timestep \leq 5~\mathrm{fs}$, not even if the KLN method is employed.
This means that the numerical artifacts do not affect the variance in total energy, which determines the $c_v$, as much as they affect the center of the corresponding distribution.

@@

Significant enhancement was achieved for a system of rigid TIP3P-water molecules \cite{Jorgensen_1983}.
Analytical derivation of a refined Hamiltonian formula for rigid bodies was greatly facilitated by considering an unsplit integration of rotations, such as in the method we have recently introduced \cite{Silveira_2017}.
In practice, we have been able to remove serious artifacts which take place when the method of Kamberaj, Low, and Neal \cite{Kamberaj_2005} or a rigid-body extension of the standard Nos\'{e}-Hoover chain thermostat \cite{Martyna_1992, Martyna_1996} is used with time steps larger than $1~\text{fs}$.
The refinement/reweighting approach described in Sec.~\ref{sec:refined_method} corrected discretization effects in computed thermodynamic properties of water for time-step sizes up to $5~\text{fs}$. This is of paramount importance as it allows one to increase the time-step size without sacrificing accuracy and, thus, to actually exploit the useful coarse-graining strategy provided by rigid-body MD.

>>>>

Fig.~\ref{fig:properties}(d) contains results for the constant-volume specific heat capacity, which
are not noticeably influenced by the discretization errors when $\timestep \leq 5~\mathrm{fs}$, not even if the KLN method is employed.
This means that the numerical artifacts do not affect the variance as much as they affect the mean of the total energy distribution.

@@

Significant enhancement was achieved for a system of rigid TIP3P-water molecules \cite{Jorgensen_1983}.
Analytical derivation of a refined Hamiltonian formula for rigid bodies was greatly facilitated by considering an unsplit integration of rotations, such as in the method we have recently introduced \cite{Silveira_2017}.
In practice, we have been able to remove serious artifacts which take place when the method of Kamberaj, Low, and Neal \cite{Kamberaj_2005} or a rigid-body extension of the standard Nos\'{e}-Hoover chain thermostat \cite{Martyna_1992, Martyna_1996} is used with time steps larger than $1~\text{fs}$.
The refinement/reweighting approach described in Sec.~\ref{sec:refined_method} corrected discretization effects in computed thermodynamic properties of water for time-step sizes up to $5~\text{fs}$. This is of paramount importance as it allows us to actually exploit the useful coarse-graining strategy provided by rigid-body MD by increasing the time-step size without sacrificing accuracy.
